Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2024-08-07 paper/main.tex
udpates

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Utilizing Expert Knowledge for Model Modification}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
	\emph{Causal Discovery} has received significant attention in the
	\emph{Directed Acyclic Graphs} (DAGs) literature, leading to the
	development of numerous automated algorithms for learning DAGs from
	data. Despite these advancements, the application of these algorithms
	in applied domains remain limited, with researchers often preferring to
	construct DAGs manually based on their expert knowledge. We hypothesize
	that the reluctance to use these algorithms arises due to several
	practical issues, such as algorithms often make obvious mistakes making
	it difficult to trust, they output the Markov Equivalence Class (MEC)
	instead of a DAG (which is often required for further analysis), and
	difficulties in choosing the appropriate algorithm and hyperparameters
	for a given dataset. Even when using these algorithms, researchers must
	rely on their expert knowledge to make modifications and orient edges
	of the learned model. Given this reliance on manual model modification,
	in this paper, we present a method inspired by \emph{specification
	search} in \emph{Structural Equation Models} (SEMs) to assist
	researchers in iteratively modifying their models to improve their fit
	to data. Specifically, we utilize implied Conditional Independence (CI)
	based model testing combined with a novel measure of partial
	association for mixed data to rank violations of implied CIs of the
	model in data. Using this ranking users can choose to make
	modifications to their models by either removing edges or adding edges
	between pair of variables while using their expert knowledge to decide
	the orientation of the edges. The ranking allows users to prioritize
	the worst offending violations first. Through empirical analysis, when
	applied to learning the model from scratch, we demonstrate that taking
	a greedy approach (i.e., fixing the worst violation first) this
	approach can get comparable performance to automated algorithms if the
	users are able to decide the edge direction correctly with an accuracy
	higher than $ 0.4 $. (TODO: Double check this value). We also show that
	a Large Language Model (LLM) can also be used for the purpose of
	deciding the edge orientation in the absence of an expert.
\end{abstract}

>>>>

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
	\emph{Causal Discovery} has received significant attention in the
	\emph{Directed Acyclic Graphs} (DAGs) literature, leading to the
	development of numerous automated algorithms for learning DAGs from
	data. Despite these advancements, the application of these algorithms
	in applied domains remain limited, with researchers often preferring to
	construct DAGs manually based on their expert knowledge. We hypothesize
	that the reluctance to use these algorithms arises due to several
	practical issues, such as algorithms often make obvious mistakes making
	it difficult to trust, they output the Markov Equivalence Class (MEC)
	instead of a DAG (which is often required for further analysis), and
	difficulties in choosing the appropriate algorithm and hyperparameters
	for a given dataset. Even if using these algorithms, researchers have
	to rely on their expert knowledge to make modifications to the learned
	model in terms of fixing any mistakes or orienting the edges before
	these models can be used for any further analysis. Given this reliance
	on manual model modification, in this paper, we present a method
	inspired by \emph{specification search} in \emph{Structural Equation
	Models} (SEMs) to assist researchers in iteratively modifying or
	constructing their models to improve their fit to data. Specifically,
	we utilize implied Conditional Independence (CI) based model testing
	combined with a novel measure of partial association for mixed data to
	rank violations of implied CIs of the model in data. Based on this
	ranking, users can then choose to make modifications to the model by
	either removing or adding edges while using their expertise to decide
	the orientation of the edge. The ranking allows users to prioritize the
	worst offending violations first. Through empirical analysis, when
	applied to constructing the model from scratch, we demonstrate that
	taking a greedy approach (i.e., fixing the worst violation first) gets
	us comparable performance to automated algorithms if the user is able
	to correctly provide the edge direction in $ 1 $ out of $ 3 $ prompts.
	We further show that in the absence of an expert, a Large Language
	Model (LLM) can also be used for getting the edge orientations.
\end{abstract}
